---
title: "Models of Neighborhood Change in San Diego, CA"
output: 
  flexdashboard::flex_dashboard:
    theme: bootstrap
    source: embed
    smart: false
runtime: shiny
---

    

```{r global, echo=F}

# PACKAGES 


# dashboard layout, widgets, and uploading

library( flexdashboard )
library( shiny )
library( rsconnect )

# data wrangling 

library( dplyr )


# formatting output

library( DT )
library( pander )
library( knitr )
library( stargazer )


# maps

library( ggmap )
library( leaflet )
library( viridis )
library( geojsonio )
library( sp )
library( sf )
library( tmap )
library( pals )
```


```{r}
#load dorling cartogram
# from local file path
#sd <- geojson_read( "C:\\Users\\swest\\Desktop\\Grad School\\Spring 2024\\PAF 516\\-sd_dorling_final.geojson", what="sp" )

#sd2 <- spTransform(sd, CRS("+init=epsg:3395"))

github.url <- "https://raw.githubusercontent.com/swest235/paf516/main/-sd_dorling_final.geojson"
sd <- geojson_read(x = github.url, what = "sp")
sd2<- spTransform(sd, CRS("+init=epsg:3395"))

sd.sf<- st_as_sf(sd2)

d<- as.data.frame(sd.sf)

#current_bbox <- st_box(geo_sf)   **Suuspect bb error was cause for object object error on dashboard - GPT suggestion

#new_bbox<- c(xmin = -117.346069, xmax = -116.815979, ymin = 32.516977, ymax = 32.912021 )  **Suuspect bb error was cause for object object error on dashboard - GPT suggestion

#sd.sf<-new_bbox  **Suuspect bb error was cause for object object error on dashboard - GPT suggestion

#bb <- st_bbox( c( xmin = -117.346069, xmax = -116.815979, 
               #   ymax = 32.912021, ymin = 32.516977 ), 
              # crs = st_crs("+init=epsg:4326"))  **neighborhoods tab looks fine without bbbox.

```




Community Demographics  
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
these.variables <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12.x", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

dd.name <- c("Percent white, non-Hispanic", 
                 "Percent black, non-Hispanic", 
                "Percent Hispanic", 
                 "Percent Native American race", 
                "Percent foreign born", 
                 "Percent speaking other language at home, age 5 plus", 
                 "Percent with high school degree or less", 
                 "Percent with 4-year college degree or more", 
                 "Percent unemployed", 
                  "Percent female labor force participation", 
                 "Percent professional employees", 
                 "Percent manufacturing employees", 
                 "Percent veteran", 
                  "Percent self-employed", 
                 "Median HH income, total", 
                  "Per capita income", 
                 "Percent in poverty, total", 
                  "Percent owner-occupied units", 
                 "Percent vacant units",
                  "Percent multi-family units", 
                 "Median rent", 
                  "Median home value", 
                 "Percent structures more than 30 years old",
                 "Percent HH in neighborhood 10 years or less", 
                 "Percent 17 and under, total", 
                  "Percent 60 and older, total",
                 "Percent 75 and older, total", 
                 "Percent currently married, not separated", 
                 "Percent widowed, divorced and separated", 
                 "Percent female-headed families with children")

name_mapping <- setNames(dd.name, these.variables)

choice_names = as.vector(name_mapping[these.variables])

# replace these with descriptive labels 
# from the data dictionary 
#temp.names <- paste0( "Variabls", these.variables )

radioButtons( inputId="demographics", 
              label = h3("Census Variables"),
              # choices = these.variables, 
              choiceNames=choice_names,
              choiceValues=these.variables,
              selected="pnhwht12")



```



Row {.tabset}
-------------------------------------


### Choropleth Map


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             sd.sf <- 
             sd.sf %>% 
             mutate( q = ntile( get(input$demographics), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Choropleth of Select Demographics: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Population Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -13071745, xmax = -12980948 ) + 
    ylim( ymin = 3811797, ymax = 3931462 )


})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$demographics ] })

x <- get_variable_x() %>% unlist()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of variable ", toupper( input$demographics ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```




Neighborhoods  
===================================== 

### Clusters  

```{r}

# define the bounding box corners 
#bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                  #ymax = 3965924, ymin = 3899074 ), 
               #crs = st_crs("+init=epsg:3395"))            ##Removed the Bounding Box and it seems to generate the close enough overview of san diego. 


# ADD YOUR CUSTOM LABELS TO THE CLUSTERS 

sd2$cluster[ sd2$cluster == "1" ] <- "Baby Boomers"
sd2$cluster[ sd2$cluster == "2" ] <- "Hipsters"
sd2$cluster[ sd2$cluster == "3" ] <- "Young Professionals"
sd2$cluster[ sd2$cluster == "4" ] <- "Working Class"
sd2$cluster[ sd2$cluster == "5" ] <- "Leisure World"
sd2$cluster[ sd2$cluster == "6" ] <- "Cultural Survivors"

renderTmap({

tmap_mode("view")
tm_basemap( "CartoDB.Positron" )
tm_shape( sd2) + 
  tm_polygons( col="cluster", palette="Accent",
               title="Community Types" ) 

})
```




NH Change 2000-2010  
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c(
  "Median Home Value 2000",
  "Median Home Value 2010",
  "Value Change 2000-2010",
  "Growth in Home Value")
button.values <- c(
  "mhv.00",
  "mhv.10",
  "mhv.change",
  "mhv.growth")

name_mapping2 <- setNames(button.labels, button.values)

choice_names2 = as.vector(name_mapping2[button.values])


 
radioButtons( inputId="home.value",
              label = h3("Home Values"),
              choiceNames=choice_names2,
              choiceValues=button.values,
              selected="mhv.00")
```




Row {.tabset}
-------------------------------------



### Median Home Values


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             sd.sf <- 
             sd.sf %>% 
             mutate( q = ntile( get(input$home.value), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Spatial Distribution of Home Values: ", toupper(input$home.value) ),
          caption = "Source: Harmonized Census Files",
          fill = "Home Value Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" )+ 
    xlim( xmin = -13071745, xmax = -12980948 ) + 
    ylim( ymin = 3811797, ymax = 3931462 ) 

})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$home.value ] }) ##by putting input$demographics here, the histogram is exactly the same between selections. The only thing that changes is the title above the graph. 

x <- get_variable_x() %>% unlist() %>% as.numeric()
x <- na.omit(x)
cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of ", toupper( input$home.value ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```  







Drivers of Change   
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.00","mhv.10","mhv.change","mhv.growth")
  
radioButtons( inputId="dv", 
              label = h3("Select Your Dependent Variable"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.change")


covariates <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12.x", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

covariate.labels <- c( "Percent white, non-Hispanic", 
                 "Percent black, non-Hispanic", "Percent Hispanic", 
                 "Percent Native American race", "Percent foreign born", 
                 "Percent speaking other language at home, age 5 plus", 
                 "Percent with high school degree or less", 
                 "Percent with 4-year college degree or more", 
                 "Percent unemployed", "Percent female labor force participation", 
                 "Percent professional employees", 
                 "Percent manufacturing employees", 
                 "Percent veteran", "Percent self-employed", 
                 "Median HH income, total", "Per capita income", 
                 "Percent in poverty, total", "Percent owner-occupied units", 
                 "Percent vacant units", "Percent multi-family units", 
                 "Median rent", "Median home value", 
                 "Percent structures more than 30 years old",
                 "Percent HH in neighborhood 10 years or less", 
                 "Percent 17 and under, total", "Percent 60 and older, total",
                 "Percent 75 and older, total", 
                 "Percent currently married, not separated", 
                 "Percent widowed, divorced and separated", 
                 "Percent female-headed families with children")
#name_mapping3 <- setNames(covariate.labels, covariates)

#choice_names3 = as.vector(covariate.labels[covariates])

checkboxGroupInput(
  inputId = "covariates",
  label = h3("Select Variables for Your Model"),
 # choices = covariates,I'm not sure why, but when i remove this the correct labels show up. 
 choiceNames = covariate.labels,
  choiceValues = covariates,
  selected = c("pnhwht12", "pprof12", "pvac12")
)








```




Row {.tabset}
-------------------------------------



### Predicting Change 

```{r, results="asis"}

# RUNNING A REGRESSION WITH USER INPUTS
#
# create a formula object
# by constructing the formula from user selections 
# as a string then casting as a formula object 
# x.s <- c("x1","x2","x3" )
# formula.text <- paste( "y", " ~ ", paste0( x.s, collapse=" + ") )
# formula.object <- as.formula( formula.text )
#
# lm( formula.object, data=d )
#
# make sure all variables are in d


  
# check boxes return vectors 
get_covariates <-  reactive({ input$covariates })


renderUI({
  
covariates <- get_covariates() 

formula.text <- paste0( input$dv, " ~ ", paste( covariates, collapse=" + " ) )
fo <- as.formula( formula.text )

m <- lm( fo, data=d )


# HTML( "<div style='width: 60%; margin: 0px auto;'>" )
HTML( 
  
  c("<br><br><br>",
  "<div type='regression' style='width: 60%; margin: 0px auto;'>",
  stargazer( m, type="html", omit.stat=c("rsq","f") ),
  "</div>",
  "<br><br><br>"
  )
  
  )


})

# HTML( reg.table )
```



### Correlation Plots 

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
usr <- par("usr"); on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex <- 0.8/strwidth(txt)

test <- cor.test(x,y)

Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
              cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
              symbols = c("***", "**", "*", ".", " "))

text(0.5, 0.5, txt, cex = 1.5 )
text(.7, .8, Signif, cex=cex, col=2)
}

panel.smooth <- function (x, y, col = par("col"), bg = NA, pch = par("pch"),
cex = 0.5, col.smooth = "red", span = 2/3, iter = 3, ...)
{
points(x, y, pch = 19, col = gray(0.7,0.2), bg = bg, cex = cex)
ok <- is.finite(x) & is.finite(y)
if (any(ok))
lines(stats::lowess(x[ok], y[ok], f = span, iter = iter),
col = col.smooth, lwd=2, ...)
}


jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

plot( x1, x2,
      pch=19, 
      col=gray(0.6, alpha = 0.2), 
      cex=0.5,  
      bty = "n",
      xlab=lab1, 
      ylab=lab2, cex.lab=1.5,
    ... )

if( draw.line==T ){ 
    ok <- is.finite(x1) & is.finite(x2)
    lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }
}


renderPlot({
pairs(d[input$covariates], upper.panel=panel.cor, lower.panel=panel.smooth)
})



```





<style>

.chart-shim { overflow: auto; }
 
table{
   border-spacing:1px;
   margin-top:30px;
   margin-bottom:30px;
   margin-left: auto;
   margin-right: auto;
   align:center} 

td{ padding: 6px 10px 6px 10px } 

th{ text-align: left; } 

</style>
